Algorithms for Cost- and Deadline-Constrained Provisioning for Scientific Workflow Ensembles in IaaS Clouds

### 2014-10-19 paper/paper.tex
applied Ewa comments

Large-scale applications expressed as scientific workflows are often grouped
into ensembles of inter-related workflows. In this paper, we address a new and
important problem concerning the efficient management of such ensembles under
budget and deadline constraints on Infrastructure-as-a-Service (IaaS) clouds.
IaaS clouds are characterized by on-demand resource provisioning capabilities
and a pay-per-use model. We discuss, develop, and assess algorithms based on
static and dynamic strategies for both task scheduling and resource
provisioning. We perform the evaluation via simulation using a set of scientific
workflow ensembles with a broad range of budget and deadline parameters, taking
into account task granularity, uncertainties in task runtime estimations,
provisioning delays, and failures. We find that the key factor determining the
performance of an algorithm is its ability to decide which workflows in an
ensemble to admit or reject for execution. Our results show that an admission
procedure based on workflow structure and estimates of task runtimes can
significantly improve the quality of solutions.

@@

There are many applications that require scientific workflow ensembles.
CyberShake~\cite{Callaghan2011}, for example, uses ensembles to generate seismic
hazard maps.  Each workflow in a CyberShake ensemble generates a hazard curve
for a particular geographic location, and several hazard curves are combined to
create a hazard map.  In a 2013 study, CyberShake was used to generate a set
of hazard maps over 286 sites that required an ensemble of 2288
workflows~\cite{CyberShake13.4}. Similarly, users of Montage~\cite{Deelman2008}
often need several workflows  with different parameters to generate a set of
image mosaics that can be combined  into a single, large mosaic. The Galactic
Plane ensemble~\cite{GalacticPlane}, which generates several mosaics of the
entire sky in different wavelengths, consists of 17 workflows,
each of which contains 900 sub-workflows.  Another ensemble example is the
Periodograms application~\cite{Vockler2011}, which searches for extrasolar
planets by detecting periodic dips in the light intensity of their host star.
Due to the large scale of the input data, this application is often split up
into multiple batches processed by different workflows. Additional workflows are
created to run the analysis using different parameters. A recent analysis of
Kepler satellite data required three ensembles of 15 workflows.

@@

Infrastructure-as-a-Service (IaaS) clouds offer the ability to provision 
resources on-demand according to a pay-per-use model. These systems 
are regarded by the scientific community as a potentially attractive source of 
low-cost computing resources~\cite{Ostermann2010, Keahey2009}. In contrast 
to clusters and grids, which typically offer best-effort quality of service, clouds 
give more flexibility in creating a controlled and managed computing environment.
Clouds provide the ability to adjust resource capacity according to the changing
demands of the application, often called auto-scaling. However, giving users 
more control also requires the development of new methods for task 
scheduling and resource provisioning. Resource management decisions required 
in cloud scenarios not only have to take into account performance-related 
metrics such as workflow makespan or resource utilization, but must also consider
budget constraints, since the resources from commercial clouds
usually have monetary costs associated with them~\cite{Durkee2010}.

@@

Policy-based approaches for scientific workloads (e.g. \cite{Marshall2010,
Kim2011,Chen2011}) also enable scaling the cloud resource pool or to extending the
capabilities of clusters using cloud-burst techniques. Our approach is different
in that we consider workflows, while policy based approaches typically consider
bags of independent tasks or unpredictable batch workloads. This enables us to
take advantage of workflow-aware heuristics that cannot be applied to
independent tasks.

@@

Scheduling optimization for multiple workflows has been also considered in previous research, 
either in the context of heterogeneous systems~\cite{Zhao2006} 
where makespan and fairness optimization is applied, 
or in the context of grids~\cite{Hirales2010} 
where multiple heuristics are evaluated with respect to the makespan of schedules they produce. 
Scheduling multiple workflows in~\cite{Duan2012,Duan2014} applies multiobjective optimization 
of time and cost under storage bandwidth and capacity constraints with the use of heuristics based on game theory. 
The approach proposed in~\cite{TolosanaCalasanz20121300} focuses on enforcement of QoS 
measured as throughput of processing by individual workflows. 
The problem of workflow admission addressed in~\cite{Zheng2013a} is similar to the one solved by our algorithms, 
but the difference is that we address the problem from the end-user perspective, 
while in their work they assume that the resource provider decides 
whether to admit or reject the workflow based on the resource estimates and the deadline constraints.

@@

The analysis of related work is summarized in Table~\ref{tab:related}, where we classify the methods based on 7 categories. 
Infrastructure type can be grid or cloud, where grid includes utility grids 
and heterogeneous computing systems where cost can be relevant (such as in~\cite{Sakellariou2007}). 
Application types of interest are bags of tasks, single workflows or multiple workflows. 
The workload can be static, where the scheduling process can plan the tasks in advance, 
or dynamic, where the algorithm has to deal with unpredictable stream of jobs arriving. 
The provisioning methods are divided into dynamic and static approaches, 
where static means that the provisioning plan is prepared prior to execution. 
To simplify, we classify all the algorithms for grid and heterogeneous systems into the static category, 
since these systems do not allow for creation of new resources on demand, 
even if some mechanisms for dynamic provisioning such as pilot jobs are available. 
The related work addresses multiple optimization objectives, 
including time, monetary cost, reliability, energy or fairness, 
and there are examples of multi-criteria optimization among them. 
Most of the scheduling approaches also consider constraints for optimization, 
and the most commonly addressed ones are budget and deadline. 
Other constraints, such as bandwidth, storage capacity or job throughput, are also present. 
Finally, we divide the algorithmic methods into heuristics, 
including various list scheduling techniques based on graph analysis, 
metaheuristics that use genetic or evolutionary algorithms, 
and the approaches using mathematical programming such as linear 
or mixed-integer programming for which ready to use solvers exist.

An analysis of the table suggests that there are many approaches to workflow scheduling 
and resource provisioning on clouds, but there is still room for improvement. 
New infrastructure types, including hybrid and federated clouds should be investigated, 
more emphasis should be put on problems with multiple workflows, 
and new methods for combining static and dynamic workloads and algorithms offer potentially interesting area of research. 
Finally, other objectives and constraints in addition to budget and execution time can be explored.

@@

To illustrate how the algorithms perform, we prepared an artificial ensemble 
of three workflows, as seen in Fig.~\ref{fig:example-ensemble}. 
The deadline is set to 6 hours and the budget to \$18. In this example we assume
that there are no VM provisioning delays, the runtime estimates are accurate, 
there are no failures and the cost of 1 VM-hour is \$1.
For dynamic algorithms we assume that the low utilization threshold ($u_l$) for deprovisioning is 50\%
and that the autoscaling parameter $v_{max} = 1$.

\begin{figure}[htb] 
  \centering
  \includegraphics[width=0.80\columnwidth]{figures/example-ensemble}
  \caption{Artificial ensemble consisting of three workflows \{a, b, c\}, with priorities \{1, 2, 3\}.
  Task labels indicate estimated execution time in minutes, 
  so e.g. task $a.70$ has estimated execution time of 70 minutes.}
  \label{fig:example-ensemble}
\end{figure}

The resulting schedule produced by DPDS algorithm is shown in Fig.~\ref{fig:algorithms-example-dpds}.
The algorithm at the beginning estimates that it needs 3 VM instances by dividing the budget by the deadline ($18 / 6 = 3$). 
Next, it schedules all the ready tasks in the priority order, so $a.70$ is started first, 
and immediately $b.70$ and $c.60$ are started, since no other tasks of workflow a are ready. 
When task $a.60$ completes, there are again other no higher priority tasks ready, 
so the next task from workflow $c$ can be started ($c.45$). Since there is no preemption,
this results in workflow $a$ being delayed by the lower priority workflow $c$.
Finally, the outcome of this dynamic scheduling policy is that none of the workflows completes within the deadline.
Actually, the tasks $b.100$, $c.65$ and $a.160$ are not finished since all the VMs 
are terminated due to running out of budget. 
The exponential score achieved by the algorithm computed according to equation~\ref{eq:score} is 0.
On the other hand, we can observe that the utilization of the VMs is kept on the level of 100\% throughout the 
whole execution time, which may be considered an advantage in some cases.

@@

The resulting schedule produced by WA-DPDS algorithm is shown in Fig.~\ref{fig:algorithms-example-wadpds}.
The schedule is generated in a similar way as in DPDS algorithm, with the one notable difference. 
After scheduling tasks $a.70$ and $b.70$, the algorithm estimates whether it has enough budget to 
admit workflow $c$. The answer is that based on the sum of task runtimes there is no room for workflow $c$,
so the whole workflow is rejected. As a consequence, task $a.100$ can start at time $t=70$ minutes
thus workflow $a$ is not delayed as in the case of DPDS. One can also observe that $VM.2$ and $VM.3$ remain idle for
one hour, which is the consequence of the conservative provisioning settings we assume in this example.  
This means that VM utilization at the level of 2/3 does not trigger deprovisioning in this case, and the maximum autoscaling
factor set to 1 prevents from provisioning of new VMs over the initial number even if utilization is 100\%.
In the resulting schedule only workflow $a$ completes, so the score is $2^{-1} = 0.5$, which is better than DPDS.

@@

The final schedule produced by the SPSS algorithm is shown in
Fig.~\ref{fig:algorithms-example-spss}. SPSS generates this schedule by
considering workflows in priority order. It first considers workflow $a$, and
develops a low-cost schedule that completes $a$ by its deadline using a minimum
number of VM-hours. This process involves assigning a deadline to each task
according to Equation~\ref{eq:spss-deadline}, and then scheduling tasks in
deadline-order on the VM that can complete each task by its deadline with the
the minimum increase in cost. Once $a$ has been scheduled, SPSS will develop an
updated low-cost schedule that includes workflow $b$. However, after adding $b$,
the algorithm will determine that the total cost of the combined schedule
for $a$ and $b$ is greater than the budget, and will reject workflow $b$ and
remove it from the schedule. It will then consider workflow $c$, and develop
a low-cost schedule that adds $c$ to the existing schedule for $a$. In order
to complete each job in $c$ by its deadline, SPSS provisions 6 additional
VM-hours, which includes adding an hour to the start time of $VM.2$ for task
$c.60$, adding an hour to the end of $VM.2$ for task $c.45$, adding an hour
to the end of $VM.3$ for $c.50$, adding a new VM, $VM.4$, for 2 hours to run
$c.65$, and finally adding another hour to the end of $VM.2$ for $c.55$. The
resulting schedule has a total cost of 16 VM-hours, or \$16, which is less
than the budget of \$18 and completes both $a$ and $c$ by the deadline.
The exponential score is then $2^{-1}+2^{-3} = 0.625$, which is better than WA-DPDS.

@@

\begin{figure}[tb] 
  \centering
  \subfloat[DPDS]{
  \includegraphics[width=0.80\columnwidth]{figures/output-test-DPDS}
  \label{fig:algorithm-example-dpds}
  }\\
  \subfloat[SPSS]{
  \includegraphics[width=0.80\columnwidth]{figures/output-test-SPSS}
  \label{fig:algorithm-example-spss}
  }
  \caption[Example schedules generated by the algorithms]{Example
  schedules generated by the algorithms. Each row is a different
  VM. Boxes are tasks colored by workflow.}
  \label{fig:examples-simulated}

@@

In order to evaluate the algorithms on a standard set of workflows, we created
randomized ensembles using workflows available from the workflow generator
gallery~\cite{WorkflowGenerator}. The
gallery contains synthetic workflows modeled using structures and parameters
that were taken from real applications. Ensembles were created using synthetic
workflows from five real applications: SIPHT, LIGO, Epigenomics, Montage and
CyberShake. For each application, workflows with 50, 100, 200, 300,
400, 500, 600, 700, 800, 900 and 1000 tasks were created. For each workflow
size, 20 different workflow instances were generated using parameters and task
runtime distributions from real workflow traces. The total collection of
synthetic workflows contains 5 applications, 11 different workflow sizes, and 20
workflow instances, for a total of 1100 synthetic workflows.

@@

\begin{figure}[t] 
    \centering
    \includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
    \caption[Histogram of workflow sizes in Pareto ensembles]{Histogram of
    workflow sizes in Pareto ensembles. Workflow size is measured in number of
    tasks.}
    \label{fig:ensemble-pareto}
\end{figure}

@@

\begin{figure*}[htb]
    \centering
    \includegraphics[width=0.9\textwidth]{run-finish-variations-test-0-output-deadlines-app-dist-selected}    
    \caption{Number of completed workflows from the ensemble depending on normalized deadline.}
    \label{fig:completed-deadlines}
\end{figure*}
\begin{figure*}[htb]
    \centering
    \includegraphics[width=0.9\textwidth]{run-finish-variations-test-0-output-budgets-app-dist-selected}    
    \caption{Number of completed workflows from the ensemble depending on normalized budget.}
    \label{fig:completed-budgets}
\end{figure*}

@@

Fig~\ref{fig:completed-budgets} shows the results based on normalized budget,
each point being the average over all deadline and random seed. What becomes
apparent is that for constant and unsorted distributions the number of completed
workflows is almost linear. This confirms that the budget constraint has the
linear influence on the results, when the influence of the deadline is
eliminated by averaging. On the other hand, for sorted distributions the
functions are convex, increasing slowly for small budgets and steeper for larger
budgets. This reflects the fact that the in sorted ensembles the largest
workflows have the highest priority, so it requires a large increment of a small
budget to fir the next workflow into the schedule.

The characteristics discussed in this section can be useful not only for
analysis of our algorithms, but also for planning scientific experiments and
understanding tradeoffs between cost, deadline and the number of work
complete~\cite{Malawski-FGCS-13}.

@@

\begin{figure*}[htb]
    \centering
%    \subfloat[CyberShake]{
    \includegraphics[width=0.6\textwidth]{run-stretching-test-output-scaling-CYBERSHAKE}
%    }\\
%    \subfloat[Montage]{
    \includegraphics[width=0.6\textwidth]{run-stretching-test-output-scaling-MONTAGE}
%    }
    \caption{Percentage of high scores achieved by each algorithm for CyberShake (top)
    and Montage (bottom) ensembles when task runtime is scaled by a constant factor. A
    scaling factor of $x$ means that the runtime of tasks in each workflow was
    multiplied by $x$.}
    \label{fig:stretching}
\end{figure*}

@@

The granularity of the workflow tasks should be discussed in relation to the
length of the cloud billing cycle. In this paper we assumed that the resources
are billed by the hour, as is the case of VM instances on Amazon EC2.
However, there are cloud providers that have higher billing frequency, e.g.
every 5 minutes in the case of
CloudSigma\footnote{\url{http://www.cloudsigma.com}} or every 1 minute in the
case of Google Compute
Engine\footnote{\url{https://cloud.google.com/pricing/compute-engine}} (after
first 10 minutes). Our results indicate that when the task granularity is close
to the length of the billing period, the SPSS produces better results, while for
tasks that are shorter than the billing period, the dynamic algorithms perform
better. This suggests that these granularity effects have to be taken into account when
selecting a cloud provider, and also when using task clustering algorithms that
group smaller tasks into larger groups~\cite{ChenD12}.

@@

Figure~\ref{fig:variances-budget} shows the ratio of simulated cost to budget. This
plot illustrates two important algorithm characteristics. First, the dynamic
algorithms very rarely exceeded the budget, even with very large errors. This
indicates that the dynamic algorithms are able to adapt to uncertainties at
runtime to ensure that the constraints are not exceeded, even when the quality
of information available to them is low. Second, unlike the dynamic
algorithms, the static algorithm frequently exceeded the budget constraint; by
large amounts in some cases. This is a result of the fact that the static
algorithm makes all of its decisions ahead of the execution and is not able to adapt
to changing circumstances.

@@

Figure~\ref{fig:failures-scores} shows the relative performance as a percentage of
the number of high scores achieved by each algorithm as the failure rate
increases. The results are similar to the ones of
Figure~\ref{fig:delays-scores}. Comparing Figure~\ref{fig:failures} to
Figs.~\ref{fig:variances} and~\ref{fig:delays} one may conclude that failures
are worse than provisioning delays and runtime estimate errors, since their
impact is larger. However, we consider higher failure rates as rare events that
suggest a significant system malfunction or invalid selection of resources.

@@

This study suggests several areas for future work. Our current approach models
data access as part of the task execution time and does not consider data
storage and transfer costs. In the future we plan to extend the application
and infrastructure model to  include the various data storage options
available on clouds. A previous experimental study \cite{Juve2010} suggests
that the data demands of scientific workflows have a large impact on not only
the execution time, but also on the cost of workflows in commercial clouds. 
In dynamic algorithms, it will be interesting to extend our
utilization-based autoscaling with more advanced workflow-aware strategies
based on feedback control.
We also plan to investigate heterogeneous environments that include multiple VM
types and cloud providers, including private and community clouds, which will
make the problem even more complex and challenging.

@@

\section*{Acknowledgments}
This work was supported by the NSF under award OCI-0943725 (STCI) and
by the EU FP7 project VPH-Share (269978).

>>>>

Large-scale applications expressed as scientific workflows are often grouped
into ensembles of inter-related workflows. In this paper, we address a new and
important problem concerning the efficient management of such ensembles under
budget and deadline constraints on Infrastructure as a Service (IaaS) clouds.
IaaS clouds are characterized by on-demand resource provisioning capabilities
and a pay-per-use model. We discuss, develop, and assess novel algorithms based on
static and dynamic strategies for both task scheduling and resource
provisioning. We perform the evaluation via simulation using a set of scientific
workflow ensembles with a broad range of budget and deadline parameters, taking
into account task granularity, uncertainties in task runtime estimations,
provisioning delays, and failures. We find that the key factor determining the
performance of an algorithm is its ability to decide which workflows in an
ensemble to admit or reject for execution. Our results show that an admission
procedure based on workflow structure and estimates of task runtimes can
significantly improve the quality of solutions.

@@

There are many applications that require scientific workflow ensembles.
CyberShake~\cite{Callaghan2011}, for example, uses ensembles to generate seismic
hazard maps.  Each workflow in a CyberShake ensemble generates a hazard curve
for a particular geographic location, and several hazard curves are combined to
create a hazard map.  In a 2013 study, CyberShake was used to generate a set
of hazard maps over 286 sites that required an ensemble of 2288
workflows~\cite{CyberShake13.4}. Similarly, users of Montage~\cite{Deelman2008}
often need several workflows  with different parameters to generate a set of
image mosaics that can be combined  into a single, large mosaic. The Galactic
Plane ensemble~\cite{GalacticPlane}, which generates several mosaics 
in different wavelengths, consists of 17 workflows,
each of which contains 900 sub-workflows.  Another ensemble example is the
Periodograms application~\cite{Vockler2011}, which searches for extrasolar
planets by detecting periodic dips in the light intensity of their host star.
Due to the large scale of the input data, this application is often split up
into multiple batches processed by different workflows. Additional workflows are
created to run the analysis using different parameters. A recent analysis of
Kepler satellite data required three ensembles of 15 workflows.

@@

Infrastructure-as-a-Service (IaaS) clouds offer the ability to provision 
resources on-demand according to a pay-per-use model. These systems 
are regarded by the scientific community as a potentially attractive source of 
low-cost computing resources~\cite{Ostermann2010, Keahey2009}. In contrast 
to clusters and grids, which typically offer best-effort quality of service, clouds 
give more flexibility in creating a controlled and managed computing environment.
Clouds provide the ability to adjust resource capacity according to the changing
demands of the application, often called auto-scaling. However, giving users 
more control also requires the development of new methods for task 
scheduling and resource provisioning. Resource management decisions required 
in cloud scenarios not only have to take into account performance-related 
metrics such as workflow makespan or resource utilization, but must also consider
budget constraints, since commercial cloud resources
usually have monetary costs associated with them~\cite{Durkee2010}.

@@

Policy-based approaches for scientific workloads (e.g. \cite{Marshall2010,
Kim2011,Chen2011}) also enable scaling the cloud resource pool or the extension of the
capabilities of clusters using cloud-burst techniques. Our approach is different
in that we consider workflows, while policy-based approaches typically consider
bags of independent tasks or unpredictable batch workloads. This enables us to
take advantage of workflow-aware heuristics that cannot be applied to
independent tasks.

@@

Scheduling optimization for multiple workflows has been also considered in previous research, 
either in the context of heterogeneous systems~\cite{Zhao2006}, 
where makespan and fairness optimization is applied, 
or in the context of grids~\cite{Hirales2010}, 
where multiple heuristics are evaluated with respect to the makespan of schedules they produce. 
Scheduling multiple workflows in~\cite{Duan2012,Duan2014} applies multiobjective optimization 
of time and cost under storage bandwidth and capacity constraints with the use of heuristics based on game theory. 
The approach proposed in~\cite{TolosanaCalasanz20121300} focuses on enforcement of QoS 
measured as throughput of processing by individual workflows. 
The problem of workflow admission addressed in~\cite{Zheng2013a} is similar to the one solved by our algorithms, 
but the difference is that we address the problem from the end-user perspective, 
while in their work they assume that the resource provider decides 
whether to admit or reject the workflow based on the resource estimates and the deadline constraints.

@@

The analysis of related work is summarized in Table~\ref{tab:related}, where we classify the methods based on 7 categories. 
The infrastructure type can be a grid or a cloud, where a grid includes utility grids 
and heterogeneous computing systems, where cost can be relevant (such as in~\cite{Sakellariou2007}). 
Application types of interest are bags of tasks, single workflows, or multiple workflows. 
The workload can be static, where the scheduling process can plan the tasks in advance, 
or dynamic, where the algorithm has to deal with unpredictable stream of jobs arriving. 
The provisioning methods are divided into dynamic and static approaches, 
where static means that the provisioning plan is prepared prior to execution. 
To simplify, we classify all the algorithms for grid and heterogeneous systems into the static category, 
since these systems do not allow for creation of new resources on demand, 
even if some mechanisms for dynamic provisioning such as pilot jobs~\cite{Sfiligoi08} are available. 
The related work addresses multiple optimization objectives, 
including time, monetary cost, reliability, energy or fairness, 
and there are examples of multi-criteria optimization among them. 
Most of the scheduling approaches also consider constraints for optimization, 
and the most commonly addressed ones are budget and deadline. 
Other constraints, such as bandwidth, storage capacity or job throughput, are also present. 
Finally, we divide the algorithmic methods into heuristics, 
including various list scheduling techniques based on graph analysis, 
metaheuristics that use genetic or evolutionary algorithms, 
and the approaches using mathematical programming such as linear 
or mixed-integer programming for which ready to use solvers exist.

An analysis of the table suggests that there are many approaches to workflow scheduling 
and resource provisioning on clouds, but there is still room for improvement. 
New infrastructure types, including hybrid and federated clouds should be investigated, 
more emphasis should be put on problems with multiple interrelated workflows, 
and new methods for combining static and dynamic workloads and algorithms offer potentially interesting areas of research. 
Finally, other objectives and constraints in addition to budget and execution time can be explored.

@@

To illustrate how the algorithms perform, we prepared an artificial ensemble 
of three workflows, as seen in Figure~\ref{fig:example-ensemble}. 
The deadline is set to 6 hours and the budget to \$18. In this example we assume
that there are no VM provisioning delays, the runtime estimates are accurate, 
there are no failures and the cost of 1 VM-hour is \$1.
For dynamic algorithms we assume that the low utilization threshold ($u_l$) for deprovisioning is 50\%
and that the autoscaling parameter $v_{max} = 1$.

\begin{figure}[htb] 
  \centering
  \includegraphics[width=0.80\columnwidth]{figures/example-ensemble}
  \caption{Artificial ensemble consisting of three workflows \{$a$, $b$, $c$\}, with priorities \{1, 2, 3\}.
  Task labels indicate estimated execution time in minutes, 
  so e.g. task $a.70$ has estimated execution time of 70 minutes.}
  \label{fig:example-ensemble}
\end{figure}

The resulting schedule produced by DPDS algorithm is shown in Figure~\ref{fig:algorithms-example-dpds}.
The algorithm at the beginning estimates that it needs 3 VM instances by dividing the budget by the deadline ($18 / 6 = 3$). 
Next, it schedules all the ready tasks in the priority order, so $a.70$ is started first, 
and immediately $b.70$ and $c.60$ are started, since no other tasks of workflow $a$ are ready. 
When task $a.60$ completes, there are again other no higher priority tasks ready, 
so the next task from workflow $c$ can be started ($c.45$). Since there is no preemption,
this results in workflow $a$ being delayed by the lower priority workflow $c$.
Finally, the outcome of this dynamic scheduling policy is that none of the workflows completes within the deadline.
Actually, tasks $b.100$, $c.65$ and $a.160$ are not finished since all the VMs 
are terminated due to running out of budget. 
The exponential score achieved by the algorithm computed according to equation~\ref{eq:score} is 0.
On the other hand, we can observe that the utilization of the VMs is kept on the level of 100\% throughout the 
whole execution time, which may be considered an advantage in some cases.

@@

The resulting schedule produced by the WA-DPDS algorithm is shown in Figure~\ref{fig:algorithms-example-wadpds}.
The schedule is generated in a similar way as in DPDS algorithm, with one notable difference. 
After scheduling tasks $a.70$ and $b.70$, the algorithm estimates whether it has enough budget to 
admit workflow $c$. Since based on the sum of task runtimes there is no room for workflow $c$,
the whole workflow is rejected. As a consequence, task $a.100$ can start at time $t=70$ minutes.
Therefore workflow $a$ is not delayed as in the case of DPDS. One can also observe that $VM.2$ and $VM.3$ remain idle for
one hour, which is the consequence of the conservative provisioning settings we assume in this example.  
This means that VM utilization at the level of 2/3 does not trigger deprovisioning in this case, and the maximum autoscaling
factor set to 1 prevents from provisioning of new VMs over the initial number even if utilization is 100\%.
In the resulting schedule only workflow $a$ completes, so the score is $2^{-1} = 0.5$, which is better than DPDS.

@@

The final schedule produced by the SPSS algorithm is shown in
Figure~\ref{fig:algorithms-example-spss}. SPSS generates this schedule by
considering workflows in priority order. It first considers workflow $a$, and
develops a low-cost schedule that completes $a$ by its deadline using a minimum
number of VM-hours. This process involves assigning a deadline to each task
according to Equation~\ref{eq:spss-deadline}, and then scheduling tasks in
deadline-order on the VM that can complete each task by its deadline with the
the minimum increase in cost. Once $a$ has been scheduled, SPSS will develop an
updated low-cost schedule that includes workflow $b$. However, after adding $b$,
the algorithm will determine that the total cost of the combined schedule
for $a$ and $b$ is greater than the budget, and will reject workflow $b$ and
remove it from the schedule. It will then consider workflow $c$, and develop
a low-cost schedule that adds $c$ to the existing schedule for $a$. In order
to complete each job in $c$ by its deadline, SPSS provisions 6 additional
VM-hours, which includes adding an hour to the start time of $VM.2$ for task
$c.60$, adding an hour to the end of $VM.2$ for task $c.45$, adding an hour
to the end of $VM.3$ for $c.50$, adding a new VM, $VM.4$, for 2 hours to run
$c.65$, and finally adding another hour to the end of $VM.2$ for $c.55$. The
resulting schedule has a total cost of 16 VM-hours, or \$16, which is less
than the budget of \$18 and completes both $a$ and $c$ by the deadline.
The exponential score is then $2^{-1}+2^{-3} = 0.625$, which is better than WA-DPDS.

@@

\begin{figure}[tb] 
  \centering
  \subfloat[DPDS]{
  \includegraphics[width=\columnwidth]{figures/output-test-DPDS}
  \label{fig:algorithm-example-dpds}
  }\\
  \subfloat[SPSS]{
  \includegraphics[width=\columnwidth]{figures/output-test-SPSS}
  \label{fig:algorithm-example-spss}
  }
  \caption[Example schedules generated by the algorithms]{Example
  schedules generated by the algorithms. Each row is a different
  VM. Boxes are tasks colored by workflow.}
  \label{fig:examples-simulated}

@@

In order to evaluate the algorithms on a standard set of workflows, we created
randomized ensembles using workflows available from the workflow generator
gallery~\cite{WorkflowGenerator,Silva2014}. The
gallery contains synthetic workflows modeled using structures and parameters
that were taken from real applications. Ensembles were created using synthetic
workflows from five real applications: SIPHT, LIGO, Epigenomics, Montage and
CyberShake. For each application, workflows with 50, 100, 200, 300,
400, 500, 600, 700, 800, 900 and 1000 tasks were created. For each workflow
size, 20 different workflow instances were generated using parameters and task
runtime distributions from real workflow traces. The total collection of
synthetic workflows contains 5 applications, 11 different workflow sizes, and 20
workflow instances, for a total of 1100 synthetic workflows.

@@

\begin{figure}[t] 
    \centering
    \includegraphics[width=0.8\columnwidth]{figures/ensemble-pareto}
    \caption[Histogram of workflow sizes in Pareto ensembles]{Histogram of
    workflow sizes in Pareto ensembles. Workflow size is measured in number of
    tasks.}
    \label{fig:ensemble-pareto}
\end{figure}

@@

\begin{figure*}[htb]
    \centering
    \includegraphics[width=\textwidth]{run-finish-variations-test-0-output-deadlines-app-dist-selected}    
    \caption{Number of completed workflows from the ensemble depending on normalized deadline.}
    \label{fig:completed-deadlines}
\end{figure*}
\begin{figure*}[htb]
    \centering
    \includegraphics[width=\textwidth]{run-finish-variations-test-0-output-budgets-app-dist-selected}    
    \caption{Number of completed workflows from the ensemble depending on normalized budget.}
    \label{fig:completed-budgets}
\end{figure*}

@@

Figure\ref{fig:completed-budgets} shows the results based on normalized budget,
each point being the average over all deadline and random seed. What becomes
apparent is that for constant and unsorted distributions the number of completed
workflows is almost linear. This confirms that the budget constraint has the
linear influence on the results, when the influence of the deadline is
eliminated by averaging. On the other hand, for sorted distributions the
functions are convex, increasing slowly for small budgets and steeper for larger
budgets. This reflects the fact that the in sorted ensembles the largest
workflows have the highest priority, so it requires a large increment of a small
budget to fit the next workflow into the schedule.

The characteristics discussed in this section can be useful not only for
analysis of our algorithms, but also for planning scientific experiments and
understanding trade-offs between cost, deadline and the number of work
complete~\cite{Malawski-FGCS-13}.

@@

\begin{figure*}[htb]
    \centering
%    \subfloat[CyberShake]{
    \includegraphics[width=0.64\textwidth]{run-stretching-test-output-scaling-CYBERSHAKE}
%    }\\
%    \subfloat[Montage]{
    \includegraphics[width=0.64\textwidth]{run-stretching-test-output-scaling-MONTAGE}
%    }
    \caption{Percentage of high scores achieved by each algorithm for CyberShake (top)
    and Montage (bottom) ensembles when task runtime is scaled by a constant factor. A
    scaling factor of $x$ means that the runtime of tasks in each workflow was
    multiplied by $x$.}
    \label{fig:stretching}
\end{figure*}

@@

The granularity of the workflow tasks should be discussed in relation to the
length of the cloud billing cycle. In this paper we assumed that the resources
are billed by the hour, as is the case of VM instances on Amazon EC2.
However, there are cloud providers that have higher billing frequency, e.g.
every 5 minutes in the case of
CloudSigma\footnote{\url{http://www.cloudsigma.com}} or every 1 minute in the
case of Google Compute
Engine\footnote{\url{https://cloud.google.com/pricing/compute-engine}} (after
first 10 minutes). Our results indicate that when the task granularity is close
to the length of the billing period,  SPSS produces better results, while for
tasks that are shorter than the billing period,  dynamic algorithms perform
better. This suggests that these granularity effects have to be taken into account when
selecting a cloud provider, and also when using task clustering algorithms that
group smaller tasks into larger groups~\cite{ChenD12}.

@@

Figure~\ref{fig:variances-budget} shows the ratio of simulated cost to budget. This
plot illustrates two important algorithm characteristics. First, the dynamic
algorithms very rarely exceeded the budget, even with very large errors. This
indicates that the dynamic algorithms are able to adapt to uncertainties at
runtime to ensure that the constraints are not exceeded, even when the quality
of information available to them is low. Second, unlike dynamic
algorithms, the static algorithm frequently exceeded the budget constraint; by
large amounts in some cases. This is a result of the fact that the static
algorithm makes all of its decisions ahead of the execution and is not able to adapt
to changing circumstances.

@@

Figure~\ref{fig:failures-scores} shows the relative performance as a percentage of
the number of high scores achieved by each algorithm as the failure rate
increases. The results are similar to the ones of
Figure~\ref{fig:delays-scores}. Comparing Figure~\ref{fig:failures} to
Figures~\ref{fig:variances} and~\ref{fig:delays} one may conclude that failures
are worse than provisioning delays and runtime estimate errors, since their
impact is larger. However, we consider higher failure rates as rare events that
suggest a significant system malfunction or invalid selection of resources.

@@

This study suggests several areas for future work. Our current approach models
data access as part of the task execution time and does not explicitly consider data
storage and transfer costs. In the future we plan to extend the application
and infrastructure model to  include the various data storage options
available on clouds. A previous experimental study \cite{Juve2010} suggests
that the data demands of scientific workflows have a large impact on not only
the execution time, but also on the cost of workflows in commercial clouds. 
In dynamic algorithms, it will be interesting to extend our
utilization-based autoscaling with more advanced workflow-aware strategies
based on feedback control.
We also plan to investigate heterogeneous environments that include multiple VM
types and cloud providers, including private and community clouds, which will
make the problem even more complex and challenging.

@@

\section*{Acknowledgments}
This work was supported by the NSF under award ACI-1148515 and
by the EU FP7 project VPH-Share (269978).
